\documentclass{report}
\usepackage{standalone}

\usepackage{minted}

\newminted{python}{frame=lines}
\newminted{bash}{frame=lines}
\usepackage{tikz}
\usetikzlibrary{patterns}
\usepackage{hyperref}
\long\def\omitthis#1{\relax}

\title{ETB Guide}
\author{Sam Owre}

\begin{document}

\maketitle
\tableofcontents

\begin{abstract}

This document is a guide to the use of the \textit{Evidential Tool Bus}
(ETB).  The ETB is used to create and maintain eveidence-based arguments,
for example, for assurance cases.  The arguments are given in a variant of
Datalog, but in addition to Datalog facts, semantic attachments may be
given in the form of \emph{wrappers} to invoke external tools and create
facts.  This guide is primarily for those wanting to run the ETB and
create their own rules and wrappers.  For more on the theoretical side of
ETB, see~\cite{DBLP:conf/vmcai/CruanesHOS13,
DBLP:conf/birthday/CruanesHMOS14}.

\end{abstract}

%\section{Introduction}

\chapter{Overview}

The Evidential Tool Bus (ETB) allows different tools communicate through a
distributed \emph{tool bus}, using \emph{rules} to generate
\emph{evidence} in support of \emph{claims}.  The idea is that the rules,
evidence, and claims should provide a convincing argument to a skeptical
observer.  But beyond support for argumentation, the rules and wrappers
can form a workflow, allowing for the argument to be developed automatically, 
and rerunning the workflow as changes require, or simply to satisfy a
skeptic.

Almost any tool may be supported by the ETB, as long as a wrapper may be
written for it.\footnote{Some tools that solely support a graphic API may be
difficult, though even in this case there are usually workarounds.}  The
ETB was originally designed with formal methods tools (provers,
modelcheckers, SMT solvers, etc.) in mind, but has been extended to
include any tool.  Note that wrappers may be written that simply collect
emails, digital signatures, review notes, etc., so the notion of a tool is
fairly broad.  Wrappers are loaded when ETB starts up.

The evidence is expected to be in digital form, though this could include
pictures, digital signatures, etc., as stand-ins for physical objects.
When claims are made, e.g., that a particular file has been proved, the
file is the evidence for the claim, and it is copied into a local ETB (Git)
repository.  The actual claim then references the file by both the name
and its SHA-1 hash, which together make a \emph{file handle}.  This means
that as changes occur to files in the system, it is easily determined
which claims need to be updated.  Note the old claims are not invalid,
only valid for a different version of the file.

Rules are written in a form of Datalog.  When an ETB node starts up, it
loads the rules and wrappers, and waits for a client to issue a
\emph{query} (also known as a \emph{goal}).  Like Datalog, the answer to a
query is expected to be finite, and all the answers are returned at once.
The body (subgoals) of a matching rule is processed from left to right,
and matching variable bindings are collected.  Any subgoal that is
satisfied creates a corresponding claim, which is stored in the claims
table.  If the body is completed, the head is considered to have
succeeded, and all the corresponding instances are added to the claims
table.

The datalog term language has been extended to include JSON forms, in
particular, numbers, strings, booleans, maps (also known as objects,
records, or dictionaries), and arrays.  A file handle is actually a map,
with \texttt{file} and \texttt{sha1} fields.

ETB can run on a single machine, or it can run in a distributed framework.
The simplest runs on a local area network, ETB daemons are run on several
machines, then connected together through XML-RPC, forming a completely
connected network.  Each node has its own rules and wrappers; the
names of the predicates associated with wrappers are shared among all the
ETB nodes.  When processing a query, if it is found to match a wrapper
predicate that is handled by another ETB node, the query is sent to it,
and the answers, along with the evidence, are returned.  This is useful,
for example, when connecting tools together that run on different
platforms.

In developing a workflow for the ETB, it is best to start by deciding what
\emph{claims} you want to make to support your arguments.  In an ideal
world, all arguments are purely deductive and supported by proof.  But
real assurance cases require such issues as hazard analysis, probabilistic
arguments, test results, human factors analysis, etc.  All of these form
evidence that, e.g., a plane is \emph{safe} to fly, and it is this kind of
argument that the ETB was developed for.  See~\ref{Rushby13:safecomp} for
more discussion of these issues.

In the following we start by walking through some simple tutorial
examples, showing how to run the ETB, and how to create rules and
wrappers.  The last few chapters are a reference guide to the
language, writing wrappers, and writing ETB clients.

%\input{etb_install}
\chapter{Getting Started with ETB}
\input{etb_tutorial}
\input{etb_setup}
%\input{etbd}
%\input{etbsh}
%\input{etb_language}
\chapter{ETB Wrappers}
\input{etb_wrappers}
%\input{etb_connect}
\chapter{ETB Clients}
\input{etb_clients}
\chapter{ETB Configuration}
\input{etb_config}

\section{}
If you installed ETB as described in \ref{install}, you should have
\texttt{etbd} and \texttt{etbsh} in your path.  In practice, the easiest
thing to do is to start from a directory with an \texttt{etb\_conf.ini}
file in it and run \texttt{etbsh} with a port, e.g.,
\begin{bashcode}
etbsh -p 12345 --etb
\end{bashcode}
The \texttt{--etb} flag indicates that the daemon should be started as a
subprocess.
The following subsections describe the ETB daemon and shell in more detail. 

\subsection{The \texttt{etbd} daemon}
The daemon the code that implements an ETB node; a workflow is defined by
interconnected ETB nodes, each with a set of rules and wrappers that
define a part of the overall workflow.

The \texttt{etbd} command starts an ETB node, which then
\begin{itemize}
\item reads a configuration file
\item loads rules and wrappers files
\item loads a file containing earlier claims and derivations
\item changes directory to the specified Git working repository
\item waits for messages from other nodes or clients
\end{itemize}
Several nodes may be running on a single computer, or across a network.
However, it is not a good idea to run two nodes at the same time that
could overwrite the same claims file.

It is important to understand that wrappers are run in the Git working
directory, so any input files need to be introduced from outside, either
from another node or from a client.  Most often \texttt{etbd} is started
in a directory that contains an \texttt{etb\_conf.ini} file, a
\texttt{wrapper} subdirectory, any number of rules files, and an
\texttt{etb\_git} subdirectory.  In this case, the defaults work except
for the \texttt{rule\_files} and \texttt{port}.
\begin{bashcode}
[etb]
port = 26532
wrappers_dir = wrappers
logic_file = logic.json.bz2
rule_files =
git_dir = etb_git
cron_period = 25
\end{bashcode}
The \texttt{etbd} also takes arguments
\begin{description}
\item[\texttt{-h}, \texttt{--help}] show a help message
\item[\texttt{-d}, \texttt{--debuglevel}] set the debug level, one of
\texttt{debug}, \texttt{info}, \texttt{warning}, \texttt{error}, or
\texttt{critical}.  Default is \texttt{info}.
\item[\texttt{--log}] a file to capture log messages; otherwise log
messages go to the terminal (i.e., stdout and stderr).
\item[\texttt{--config}, \texttt{-cf}] the config file to use
\item[\texttt{--port}, \texttt{-p}] the port to use
\end{description}
These arguments override the corresponding configuration file arguments.

\subsection{The \texttt{etbsh} shell}
The \texttt{etbsh} is a shell client that allows for interaction with an
ETB network.  It can also start up an \texttt{etbd} daemon and connect to
it, as shown above.  The \texttt{etbsh} takes arguments
\begin{description}
\item[\texttt{-h}, \texttt{--help}] show a help message
\item[\texttt{--host}] an ETB host URL, default localhost
\item[\texttt{--port}, \texttt{-p}] the port to use
\item[\texttt{--name}, \texttt{-n}] an etb name (used in proxying)
\item[\texttt{--batch}] run in batch mode, using stdin or a SCRIPT
\item[\texttt{--etb}] run \texttt{etbd}, from the specified directory if
given, else the current directory
\item[SCRIPT] is an \texttt{etbsh} load file, see below.
\end{description}

The \texttt{etbsh} shell takes a number of commands:
\begin{description}
\item[\texttt{help} [\emph{command}]]{Help for the given \emph{command},
or a summary of all commands if no \emph{command} given.}
\end{description}


\section{ETB Language}
ETB is best viewed as an extension of Datalog.  An ETB program consists of
\emph{sentences}, and the main results are \emph{claims} that are kept in
the claims table.  In order to increase the expressiveness, ETB uses a
JSON-like extension.

\subsection{Lexical}
The basic lexical tokens are identifiers, strings, and numbers.
Identifiers are any sequence of characters, except for space, \texttt{=},
colon, period, question mark, comma, quote marks (single, double, and
back), tilde (\texttt{~}, percent \texttt{\%} and various brackets
\texttt{[]()\{\}}.  Alternatively, any sequence of characters within
single quotes forms an identifier, with the usual backslash escape to
include single quotes or backslash.  Strings are any sequence of
characters within double quotes, with the backslash convention.  Numbers
are integers or floating point, with 

Identifiers are either variables or constants.  An identifier that starts
with an upper case letter is a variable, otherwise it is a constant.
Strings are enclosed in double quotes.



\bibliographystyle{plain}
\bibliography{etb}

\end{document}

%% Local Variables:
%% TeX-command-extra-options: "-shell-escape"
%% End:
